---
title: "DuckDB"
---


# Querying Parquet with Different Engines

## Introduction to DuckDB

![](https://duckdb.org/images/duckdb-circle.svg){width=25%}

::: {.incremental}
- **Analytical SQL database system**
  - Embedded database (like SQLite)
  - Column oriented
  - In-process query execution

- **Key features:**
  - Direct Parquet querying
  - Vectorized query execution
  - Parallel processing
  - Zero-copy integration with arrow
:::

:::{.notes}
The zero-copy integration with arrow is because DuckDB uses basically the same format for it's own internal representation.
:::

## DuckDB

```{r}
#| eval: false
library(duckdb)

con <- dbConnect(duckdb())

# Register a Parquet file as a virtual table
dbExecute(con, "CREATE VIEW pums AS SELECT * 
                FROM read_parquet('CA_person_2021.parquet')")

# Run our query
dbGetQuery(con, "
  SELECT SUM(JWMNP * PWGTP)/SUM(PWGTP) as avg_commute_time,
         COUNT(*) as count
  FROM pums
  WHERE AGEP >= 16
")

dbDisconnect(con, shutdown = TRUE)
```

## duckplyr

```{r}
#| eval: false
library(duckplyr)

# Read data with Arrow
pums_data <- read_file_duckdb(
  "CA_person_2021.parquet", 
  "read_parquet"
)

# Use duckplyr to optimize dplyr operations
pums_data |>
  filter(AGEP >= 16) |>
  summarize(
    mean_commute_time = sum(JWMNP * PWGTP, na.rm = TRUE) /
      sum(PWGTP),
    count = n()
  ) |>
  collect()
```

:::{.notes}
duckplyr is a drop-in replacement for dplyr, using duckdb as a backend
:::

## data.table

```{r}
#| eval: false
library(arrow)
library(data.table)

# Read Parquet file with Arrow
pums_data <- read_parquet("CA_person_2021.parquet")

# Convert to data.table
pums_dt <- as.data.table(pums_data)

# data.table query
pums_dt[AGEP >= 16,
  .(avg_commute_time = sum(JWMNP * PWGTP, na.rm = TRUE) / sum(PWGTP), 
    count = .N)]
```

## Arrow Query Execution

```{r}
#| eval: false
# Create an Arrow Dataset
pums_ds <- open_dataset("pums_dataset_dir/")

# Execute query with Arrow
result <- pums_ds |>
  filter(AGEP >= 16) |>
  group_by(ST) |>
  summarize(
    avg_commute_time = mean(JWMNP, na.rm = TRUE),
    count = n()
  ) |>
  arrange(desc(avg_commute_time)) |>
  collect()
```

## Demo: Seamless Integration Arrow ↔ DuckDB

```{r}
#| eval: false
df <- read_parquet("CA_person_2021.parquet")

# Use dplyr verbs with arrow tables
df |>
  filter(AGEP >= 16) |>
  to_duckdb() |>
  summarize(
    mean_commute_time = sum(JWMNP * PWGTP, na.rm = TRUE) /
      sum(PWGTP),
    count = n()
  ) 
```
